---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# multigait

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/jhuwit/multigait/graph/badge.svg)](https://app.codecov.io/gh/jhuwit/multigait)
[![R-CMD-check](https://github.com/jhuwit/multigait/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jhuwit/multigait/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`multigait` is an R interface to [MultiGait](https://github.com/DMegaritis/multigait),
a Python pipeline for wrist-worn IMU gait analysis. It provides both complete
predefined pipelines and modular functions for gait-sequence detection, initial
contacts, cadence, stride length, walking speed, and walking-bout aggregation.

## Installation

``` r
remotes::install_github("jhuwit/multigait")
```

The package calls `py_require_multigait()` when it loads. With current
`reticulate`, MultiGait is installed into a managed Python environment the first
time it is needed. Confirm the setup with `have_multigait()`.

`as_multigait_data()` standardizes compatible `X/Y/Z` inputs with `actibase`,
renames them for MultiGait, and by default converts values that look like g to
m/s². Set `scale = "none"` when units are already known, or `scale = "error"`
to require an explicit unit decision.

## Full pipeline

```{r example, eval = FALSE}
library(multigait)
raw <- actiread::acti_read_gt3x(actiread::acti_example_gt3x())
wrist_imu <- data.frame(
  acc_x = raw$X * 9.80665,
  acc_y = raw$Y * 9.80665,
  acc_z = raw$Z * 9.80665
)
recording <- as_multigait_data(
  wrist_imu,
  sample_rate = 100,
  participant_metadata = list(height_m = 1.70, arm_length_cm = 65)
)
results <- run_multimorbidity_impaired_pipeline(recording)
results$aggregated
```

`run_healthy_comorbidity_pipeline()` provides the alternative configuration.

## Modular analysis

```{r modular, eval = FALSE}
contacts <- detect_initial_contacts(
  wrist_imu,
  initial_contact_detector("zijlstra", version = "wrist")
)
cadence <- calculate_cadence(wrist_imu, contacts)
```

See `vignette("multigait")` for an end-to-end workflow and information about
required sensor fields and participant metadata.
