---
title: "Using MultiGait from R"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using MultiGait from R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`multigait` exposes the Python [MultiGait](https://github.com/DMegaritis/multigait)
pipeline through `reticulate`. MultiGait estimates gait sequences, initial contacts,
cadence, stride length, walking speed, and walking-bout digital mobility outcomes
from wrist-worn IMU data.

`as_multigait_data()` uses `actibase::acti_standardize_data()` by default for
compatible `X/Y/Z` data. It then guesses whether acceleration is in g using the
typical three-axis magnitude and, when appropriate, converts it to m/s^2. Set
`scale = "none"` to preserve known units or `scale = "error"` to stop when a
g-to-m/s^2 conversion would otherwise be inferred.

## Python setup

On package load, `multigait` calls `py_require_multigait()`. With reticulate 1.41
or later, the first function that needs Python creates a managed environment with
the `multigait` Python package. To confirm availability:

```{r setup, eval = FALSE}
library(multigait)
have_multigait()
```

Some macOS installations build `numba` from source. If that build cannot find
OpenMP headers, start R with `NUMBA_DISABLE_OPENMP=true` or install an OpenMP
toolchain before the first managed-environment resolution.

## Run a complete pipeline

Prepare one sample per row of body-frame accelerometer data and provide the
sampling rate and participant information. The exact sensor-axis requirements
depend on the selected upstream algorithm.

```{r full-pipeline, eval = FALSE}
raw <- actiread::acti_read_gt3x(actiread::acti_example_gt3x())
imu <- data.frame(
  acc_x = raw$X * 9.80665,
  acc_y = raw$Y * 9.80665,
  acc_z = raw$Z * 9.80665
)
recording <- as_multigait_data(
  imu,
  sample_rate = 100,
  participant_metadata = list(height_m = 1.70, arm_length_cm = 65),
  recording_metadata = list()
)

results <- run_multimorbidity_impaired_pipeline(recording)
results$aggregated
results$per_walking_bout
```

When `actiread` is installed, use one of its readers to load its bundled
wearable examples or your own supported files, then select/rename the required
body-frame acceleration columns before calling `as_multigait_data()`. This
keeps the source recordings in the package designed to distribute them rather
than copying a large fixture into `multigait`.

Use `run_healthy_comorbidity_pipeline()` for the alternative predefined
configuration. Each result includes the Python pipeline object and R data frames
for detected gait sequences, contacts, per-second and per-stride estimates,
walking bouts, and recording-level outcomes.

## Adapt individual stages

The components are independently available when a study needs a custom pipeline.

```{r components, eval = FALSE}
detector <- gait_sequence_detector("ionescu", version = "wrist")
gait_sequences <- detect_gait_sequences(imu, detector)

contacts <- detect_initial_contacts(
  imu,
  initial_contact_detector("zijlstra", version = "wrist")
)
cadence <- calculate_cadence(imu, contacts)
stride_length <- calculate_stride_length(
  imu, contacts,
  stride_length_calculator("weinberg", version = "wrist"),
  participant_metadata = list(arm_length_cm = 65)
)
walking_speed <- calculate_walking_speed(imu, cadence, stride_length, contacts)
```

The wrappers intentionally return ordinary R data frames. Pass a Python object
from `multigait_module()` directly when an upstream feature is not yet exposed.
