---
title: "Using MultiGait from R"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using MultiGait from R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{actiread}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`multigait` exposes the Python [MultiGait](https://github.com/DMegaritis/multigait)
pipeline through `reticulate`. MultiGait estimates gait sequences, initial contacts,
cadence, stride length, walking speed, and walking-bout digital mobility outcomes
from wrist-worn IMU data.

`as_multigait_data()` uses `actibase::acti_standardize_data()` by default for
compatible `X/Y/Z` data. It then guesses whether acceleration is in g using the
typical three-axis magnitude and, when appropriate, converts it to m/s^2. Set
`scale = "none"` to preserve known units or `scale = "error"` to stop when a
g-to-m/s^2 conversion would otherwise be inferred.

## Python setup

On package load, `multigait` calls `py_require_multigait()`. This vignette
checks the required R and Python dependencies before running any analysis.

```{r setup}
library(multigait)
stopifnot(
  requireNamespace("actibase", quietly = TRUE),
  requireNamespace("actiread", quietly = TRUE),
  requireNamespace("reticulate", quietly = TRUE),
  have_multigait()
)
```

```{r dependency-details}
configuration <- reticulate::py_config()
stopifnot(reticulate::py_module_available("multigait"))
configuration$python
```

Some macOS installations build `numba` from source. If that build cannot find
OpenMP headers, start R with `NUMBA_DISABLE_OPENMP=true` or install an OpenMP
toolchain before the first managed-environment resolution.

## Run a complete pipeline

The bundled `wrist_imu` data were prepared from the `actiread` GT3X fixture.
They already contain acceleration in m/s^2 with names accepted by MultiGait.

```{r full-pipeline}
imu <- wrist_imu[c("acc_x", "acc_y", "acc_z")]
recording <- as_multigait_data(
  imu,
  sample_rate = 100,
  participant_metadata = list(height_m = 1.70, arm_length_cm = 65),
  recording_metadata = list()
)

results <- run_multimorbidity_impaired_pipeline(recording)
stopifnot(
  inherits(results$aggregated, "data.frame"),
  inherits(results$per_walking_bout, "data.frame"),
  nrow(results$aggregated) == 1L,
  nrow(results$per_walking_bout) > 0L
)
results$aggregated[, seq_len(min(6L, ncol(results$aggregated))), drop = FALSE]
```

Use `run_healthy_comorbidity_pipeline()` for the alternative predefined
configuration. Each result includes the Python pipeline object and R data frames
for detected gait sequences, contacts, per-second and per-stride estimates,
walking bouts, and recording-level outcomes.

## Adapt individual stages

The components are independently available when a study needs a custom pipeline.
The modular functions operate on body-frame names, so this example makes that
conversion explicitly.

```{r components}
component_imu <- wrist_imu[c("acc_x", "acc_y", "acc_z")]
names(component_imu) <- c("acc_is", "acc_ml", "acc_pa")

gait_sequences <- detect_gait_sequences(
  component_imu, gait_sequence_detector("ionescu", version = "wrist")
)
contacts <- detect_initial_contacts(
  component_imu, initial_contact_detector("zijlstra", version = "wrist")
)
cadence <- calculate_cadence(component_imu, contacts, sample_rate = 100)
stride_length <- calculate_stride_length(
  component_imu, contacts,
  stride_length_calculator("weinberg", version = "wrist"),
  sample_rate = 100,
  participant_metadata = list(arm_length_cm = 65)
)
walking_speed <- calculate_walking_speed(
  component_imu, cadence, stride_length, contacts, sample_rate = 100
)
stopifnot(
  inherits(gait_sequences, "data.frame"),
  inherits(contacts, "data.frame"),
  inherits(cadence, "data.frame"),
  inherits(stride_length, "data.frame"),
  inherits(walking_speed, "data.frame"),
  nrow(contacts) > 0L,
  nrow(cadence) > 0L
)
head(cadence)
```

The wrappers intentionally return ordinary R data frames. Pass a Python object
from `multigait_module()` directly when an upstream feature is not yet exposed.
